---
title: "stackedbarCompositionClones.Rmd"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

# Stacked Bar plots of whole-tumour clone composition 

```{r}

library(ggplot2)
library(readxl)
library(ggpubr)
library(dplyr)

```

## read data
```{r}
data_path <- "../Source data for figures/Source_Data_Figure_2/SourceData_Fig2c_part1.xlsx"
data_fig2ci <- read_excel(path = data_path, sheet = "SourceDataFig2ci")
data_fig2cii <- read_excel(path = data_path, sheet = "SourceDataFig2cii")
data_fig2ciii <- read_excel(path = data_path, sheet = "SourceDataFig2ciii")
```

## plotting function
```{r}
stacked_bar_plot <- function(data) {
  data_sorted <- arrange(data, Subclone, desc(CCF))
  data_sorted$SimID <- factor(data_sorted$SimID, levels = unique(data_sorted[which(data_sorted$Subclone == "loss_3p/VHL"), ]$SimID))
  
  # most of the cases (under some conditions, fewer levels need to be used for proper plotting)
  levels = c("other", "subclone_5", "subclone_4", "subclone_3", "subclone_2", "subclone_1", "loss_3p/VHL")
  fill_values = c("grey", "saddlebrown", "steelblue", "chocolate", "palegreen4", "yellow", "blueviolet")
  
  theme_set(theme_grey())
  p <- ggplot(data_sorted, aes( fill=factor(Subclone, levels=levels),   y=CCF, x=as.factor(SimID)   )   ) + 
      geom_bar(position="stack", stat="identity") +
      xlab("Model replicas")+ ylab("Whole tumour CCF (tumour level)") +
      theme(axis.text.y=element_blank(),
            axis.ticks.y=element_blank()) +
      coord_flip() +
      labs(fill = "Subclone") +
      scale_fill_manual(values = fill_values) +
      scale_x_discrete(limits = rev(levels(data_sorted$SimID))) +
    theme(
      axis.text.x = element_text(size = 6),
      axis.title.x = element_text(size = 8),
      axis.title.y = element_text(size = 8),
      legend.text = element_text(size = 6),
      legend.title = element_text(size = 8)
    )
  
  print (p)
}
```

## Figure 2ci (part 1 : whole-tumour clone composition)
```{r}
stacked_bar_plot(data_fig2ci)
```
## Figure 2cii (part 1 : whole-tumour clone composition)
```{r}
stacked_bar_plot(data_fig2cii)
```
## Figure 2ciii (part 1 : whole-tumour clone composition)
```{r}
stacked_bar_plot(data_fig2ciii)
```

## Other Supplemental figure panels that can be generated using the plotting fuction above:
* Extended Data Figure 1c (File: SourceData_ExtendedData_Fig1c_part1.xlsx)
* Extended Data Figure 2c-d (File: SourceData_ExtendedData_Fig2cd.xlsx)
* Extended Data Figure 4 (File: SourceData_ExtendedData_Fig4a_part1.xlsx; SourceData_ExtendedData_Fig4b_part1.xlsx)
